---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
activity <- read.csv('activity.csv')
```

## What is mean total number of steps taken per day?
```{r}
daily_steps <- aggregate(steps~date,activity, FUN=sum)
mean_steps <- mean(daily_steps$steps, na.rm = TRUE)
median_steps <- median(daily_steps$steps, na.rm = TRUE)
```
The mean total number of steps taken per day is `r mean_steps`. The median total number of steps taken per day is `r median_steps`. 
The histogram of the total number of steps taken each day is as follows.
```{r}
hist(daily_steps$steps, xlab="Daily Steps", main="Histogram of daily steps")
```

## What is the average daily activity pattern?
The average daily pattern plot is shown below.
```{r}
daily_pattern <- aggregate(steps~interval,activity, FUN=mean)
plot(daily_pattern$interval,daily_pattern$steps, type ="l",xlab="Interval",ylab="Mean daily steps")
max_daily_pattern <- max(daily_pattern$steps)
```
The maximum number of steps taken on average during an interval is `r max_daily_pattern`.

## Imputing missing values
```{r}
na_rows_num <- nrow(activity[activity$steps=='NA', ])
```
The number of rows with missing values is `r na_rows_num`.
The imputing strategy is using the mean value of the particular interval and using it to populate the corresponding activity value.
```{r}
for (i in 1:nrow(activity)){
  if (is.na(activity$steps[i])){
    x = which(activity[i,3] == daily_pattern$interval)
    if (length(x)==0){
      activity$steps[i] = 0
    }
    else
    {
      activity[i,1] = floor(daily_pattern[x,2])
    }
  }
}
daily_steps_new <- aggregate(steps~date,activity, FUN=sum)
mean_steps_updated <- mean(daily_steps_new$steps)
median_steps_updated <- median(daily_steps_new$steps)
```
The updated mean number of daily steps is `r mean_steps_updated` and the updated median number of daily steps is `r median_steps_updated`.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(chron)
activity$day_type = ifelse(is.weekend(as.Date(activity$date)),"weekend","weekday")
library(lattice)
activity <- transform(activity, day_type = factor(day_type))
daily_pattern_new <- aggregate(steps~day_type*interval,activity, FUN=mean)
xyplot(steps ~ interval | day_type, data = daily_pattern_new, layout=c(1,2), type = "l")
```

```{r}

```